% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math} % this also loads fontspec
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else
  % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\author{}
\date{}

\begin{document}

\hypertarget{trust-dynamics-in-multi-agent-strategic-interaction-a-simulation-study-of-bayesian-reputation-systems-in-8-player-limit-texas-holdem}{%
\section{Trust Dynamics in Multi-Agent Strategic Interaction: A
Simulation Study of Bayesian Reputation Systems in 8-Player Limit Texas
Hold'em}\label{trust-dynamics-in-multi-agent-strategic-interaction-a-simulation-study-of-bayesian-reputation-systems-in-8-player-limit-texas-holdem}}

\textbf{Rachit Agrawal} Polygence Research Project, 2025--2026

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\hypertarget{abstract}{%
\subsection{Abstract}\label{abstract}}

Reputation systems that infer trustworthiness from observed behavior are
ubiquitous in online marketplaces, social networks, and financial
platforms. Yet it remains unclear whether such systems inherently reward
cooperative behavior or create exploitable dynamics that favor
aggression. We investigate this question using a controlled multi-agent
simulation of 8-player Limit Texas Hold'em, where eight strategically
distinct agents --- five static and three adaptive --- interact over
hundreds of thousands of hands while maintaining Bayesian posterior
beliefs about each other's types. Each agent updates a probability
distribution over opponent archetypes after every observed action,
producing a rich dataset of trust evolution under repeated strategic
interaction.

We report five principal findings across three independent agent
architectures. First, we observe a strong anticorrelation between trust
and economic performance (Pearson r = −0.752 across 5 seeds × 10 000
hands of frozen rule-based agents): agents perceived as most trustworthy
by their opponents accumulate the least wealth, while the least-trusted
aggressive agent dominates through fold equity rather than superior
play. Second, we demonstrate a hard classification ceiling --- only 3 to
4 of 8 archetypes are reliably identifiable through behavioral
observation, with mathematically similar types forming an
indistinguishable cluster regardless of sample size. Third, we test
whether the trap is robust to \emph{adaptation} by allowing each agent
to optimize its own decision parameters via a per-cycle hill-climber
within an archetype-shaped bound box: the trust--profit anticorrelation
softens by Δr = +0.116 (consistent across all five seeds) but does not
close, settling at r = −0.637. Fourth, we replace the rule-based and
adaptive agents with LLM-driven agents
(\texttt{claude-haiku-4-5-20251001}, 8 independent personality
role-players) and observe the trap softens an additional Δr = +0.127 to
r = −0.510 --- the increment from Phase 1 to Phase 2 and the increment
from Phase 2 to Phase 3 are essentially equal. Fifth, the four
behavioral dimensions targeted by the metrics framework (Context
Sensitivity, Opponent Adaptation, Non-Stationarity, Strategic
Unpredictability) all \emph{fail} to meet their thresholds under LLM
agents --- three move backward --- demonstrating that LLMs faithfully
role-play archetypes but do not spontaneously develop
opponent-conditional, time-varying, or unpredictable strategy.

These results have implications for the design of reputation systems in
adversarial environments: observation-based trust inference, in the
absence of external enforcement mechanisms, systematically advantages
exploitative strategies over cooperative ones; bounded numerical
adaptation chips at the trap but does not close it; and LLM-driven
role-playing produces a comparable additional reduction without
unlocking the higher-order strategic dimensions that would distinguish
``agents that reason about reputation'' from ``agents that play their
characters.''

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\hypertarget{introduction}{%
\subsection{1. Introduction}\label{introduction}}

\hypertarget{motivation}{%
\subsubsection{1.1 Motivation}\label{motivation}}

Trust is a foundational element of social and economic interaction. In
contexts where agents cannot directly verify each other's intentions ---
online commerce, anonymous networks, international diplomacy --- trust
must be inferred from observed behavior over repeated interactions. This
inference process is formalized in many real-world systems as a
reputation mechanism: a buyer rates a seller, a lender evaluates a
borrower's credit history, or a social network ranks content creators by
engagement patterns.

A critical but under-examined question is whether these
observation-based reputation systems inherently reward the behavior they
are designed to detect. If trustworthy behavior leads to higher trust
scores, and higher trust scores lead to better outcomes, then the system
functions as intended. But if trustworthy behavior is \emph{costly} ---
if it creates predictability that opponents can exploit --- then
reputation systems may inadvertently penalize the very behavior they
incentivize.

This paper investigates this question in a controlled multi-agent
setting: 8-player Limit Texas Hold'em poker. Poker is a canonical domain
for studying strategic interaction under incomplete information (Von
Neumann \& Morgenstern, 1944; Nash, 1950). It combines stochastic
elements (card deals), hidden information (hole cards), sequential
decision-making (betting rounds), and social signaling (bluffing and
folding patterns). Critically, poker requires agents to form beliefs
about opponents' strategies --- making it a natural testbed for trust
dynamics.

\hypertarget{research-question}{%
\subsubsection{1.2 Research Question}\label{research-question}}

We ask: \textbf{In repeated strategic interactions with incomplete
information, does observation-based trust inference inherently reward
exploitation over cooperation, and is this dynamic a structural property
of the interaction system or an artifact of agent design?}

To answer this, we construct a simulation with three design choices that
distinguish it from prior work:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  \textbf{Diverse strategic archetypes.} Rather than studying a single
  pair of strategies (as in iterated Prisoner's Dilemma), we deploy
  eight distinct agents drawn from game theory and behavioral economics
  --- ranging from Nash equilibrium play to grudge-based retaliation ---
  creating a richer ecology of interaction.
\item
  \textbf{Bayesian trust model.} Every agent maintains a full posterior
  distribution over every other agent's type, updated after every
  observed action. This produces a continuous, evolving measure of trust
  rather than a binary cooperate/defect signal.
\item
  \textbf{Two-phase test of robustness.} Phase 1 implements the agents
  as hand-coded rule-based systems with frozen parameters; Phase 2
  retains those agents but lets each one tune its own parameters via a
  per-cycle hill-climbing optimizer within an archetype-shaped bound
  box. The first phase characterizes the trap; the second tests whether
  bounded numerical adaptation can escape it.
\end{enumerate}

\hypertarget{summary-of-contributions}{%
\subsubsection{1.3 Summary of
Contributions}\label{summary-of-contributions}}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  \textbf{Trust--profit anticorrelation.} We demonstrate a Pearson
  correlation of r = −0.752 (Phase 1, frozen rule-based agents), r =
  −0.637 (Phase 2, agents allowed to optimize their own parameters), and
  r = −0.510 (Phase 3, LLM personality role-players) between agents'
  mean trust scores and final chip counts, establishing that
  trustworthiness is economically costly in this setting and that
  successive tiers of agent intelligence each chip \textasciitilde0.12
  off the trap without closing it.
\item
  \textbf{Aggression dominance via avoidance.} The most aggressive agent
  (Firestorm) achieves the highest economic outcome not through superior
  hand evaluation but through 87.1\% fold equity --- opponents surrender
  pots rather than engage, producing a dynamic analogous to aggressive
  pricing in oligopolistic markets.
\item
  \textbf{Classification ceiling.} We prove that behavioral observation
  alone cannot distinguish more than 3--4 of 8 archetypes, regardless of
  sample size, due to parameter overlap in the Bayesian likelihood
  model. This establishes a fundamental limit on reputation system
  accuracy.
\item
  \textbf{Bounded optimization is insufficient.} Even when each agent is
  given a per-cycle hill-climber that tunes its own parameters within an
  archetype-shaped bound box (5 seeds × 10 000 hands), the trap softens
  by only Δr = +0.116 across all five seeds, and Opponent Adaptation
  stays at OA = 0.0003 (essentially zero). This isolates the
  \emph{adaptive} component from the \emph{opponent-modeling} component.
\item
  \textbf{LLMs role-play archetypes but do not invent strategy.}
  Replacing the rule-based agents with eight independent LLM personality
  role-players (Phase 3) further softens the trap by Δr = +0.127
  (matching the Phase 1→2 step) but produces \emph{no} increase in
  opponent-conditional, time-varying, or unpredictable behavior. Three
  of the four behavioral dimension targets actually move backward. The
  diagnostic finding: LLMs faithfully follow descriptive personality
  specs but do not spontaneously generate higher-order strategic
  reasoning that those specs do not explicitly require.
\end{enumerate}

\hypertarget{paper-organization}{%
\subsubsection{1.4 Paper Organization}\label{paper-organization}}

Section 2 reviews related work in trust modeling, multi-agent systems,
and poker AI. Section 3 describes the simulation methodology: game
engine, agent archetypes, Bayesian trust model, and the
bounded-optimization pipeline that defines Phase 2. Section 4 details
the experimental setup. Section 5 presents results from both phases.
Section 6 discusses implications, limitations, and connections to
real-world reputation systems. Section 7 concludes and outlines future
work.

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\hypertarget{related-work}{%
\subsection{2. Related Work}\label{related-work}}

\hypertarget{trust-and-cooperation-in-repeated-games}{%
\subsubsection{2.1 Trust and Cooperation in Repeated
Games}\label{trust-and-cooperation-in-repeated-games}}

The study of trust in strategic settings originates with the iterated
Prisoner's Dilemma (IPD). Axelrod's (1984) tournament demonstrated that
\emph{Tit-for-Tat} --- a strategy that cooperates first, retaliates
against defection, and forgives after punishment --- outperforms both
unconditional cooperation and unconditional defection across diverse
opponent populations. Axelrod identified four properties of successful
strategies: niceness (cooperate first), retaliation (punish defection),
forgiveness (return to cooperation), and clarity (be easy to model).

Nicky Case's interactive simulation \emph{The Evolution of Trust} (2017)
popularized these findings, mapping IPD strategies to intuitive
character archetypes: the Always Cooperator, the Always Cheater, the
Copycat (Tit-for-Tat), the Grudger, and the Detective. Our eight
archetypes are directly inspired by this mapping but extend it to a
richer action space (five actions vs.~two) and a multi-agent setting
(eight players vs.~pairwise).

A key limitation of IPD-based trust research is the binary action space:
cooperate or defect. Real-world trust decisions involve a continuum of
commitment levels, partial signals, and noisy observations. Our poker
setting naturally provides this richer structure --- agents can fold,
check, call, bet, or raise, each conveying different information about
hand strength and strategic intent.

\hypertarget{bayesian-opponent-modeling}{%
\subsubsection{2.2 Bayesian Opponent
Modeling}\label{bayesian-opponent-modeling}}

Bayesian approaches to opponent modeling maintain a probability
distribution over possible opponent types and update it as new
observations arrive. Harsanyi (1967) formalized this as games of
incomplete information, where players hold beliefs about opponents'
types drawn from a common prior. In practice, Bayesian opponent modeling
has been applied in poker (Southey et al., 2005), negotiation (Zeng \&
Sycara, 1998), and security games (Pita et al., 2010).

Our trust model follows this tradition: each agent maintains an
8-dimensional posterior over opponent archetypes, updated via
precomputed likelihood tables after every observed action. We extend
standard approaches with two mechanisms: \emph{trembling-hand noise} (ε
= 0.05), which prevents posterior collapse by ensuring no observation
has zero likelihood under any type, and \emph{exponential decay} (λ =
0.95), which down-weights older observations to allow the model to track
non-stationary opponents.

\hypertarget{multi-agent-systems-and-emergent-behavior}{%
\subsubsection{2.3 Multi-Agent Systems and Emergent
Behavior}\label{multi-agent-systems-and-emergent-behavior}}

The study of emergent dynamics in multi-agent systems has a rich
history. Shoham and Leyton-Brown (2008) provide a comprehensive
framework for multi-agent interaction, while Sandholm and Crites (1996)
demonstrated that cooperative behavior can emerge from self-interested
agents through repeated interaction. More recently, Leibo et al.~(2017)
showed that the balance between cooperative and competitive behavior in
multi-agent reinforcement learning is sensitive to environmental
parameters.

Our work differs from the reinforcement learning tradition in that our
agents do not learn to optimize a reward signal. Instead, they follow
fixed (Phase 1) or learned-but-static (Phase 2) behavioral policies
while maintaining evolving beliefs about opponents. This separation of
\emph{behavior} from \emph{belief} allows us to isolate the dynamics of
trust inference from the dynamics of strategy learning.

\hypertarget{poker-as-a-research-domain}{%
\subsubsection{2.4 Poker as a Research
Domain}\label{poker-as-a-research-domain}}

Poker has served as a benchmark for AI research since the early work of
Waterman (1970) and Findler (1977). The field achieved a milestone with
Bowling et al.'s (2015) essentially perfect solution for heads-up Limit
Texas Hold'em using counterfactual regret minimization (CFR). Moravčík
et al.~(2017) and Brown and Sandholm (2018, 2019) extended these results
to no-limit variants with DeepStack and Pluribus, respectively.

Our work does not aim to solve poker or produce optimal play. Instead,
we use poker as an \emph{environment} for studying trust dynamics ---
the game mechanics are a means to generate repeated strategic
interactions with incomplete information, bluffing incentives, and
observable behavioral signals. This positions our work closer to
Billings et al.'s (1998) opponent modeling research than to the
game-solving tradition.

\hypertarget{reputation-systems}{%
\subsubsection{2.5 Reputation Systems}\label{reputation-systems}}

The design of reputation systems for online platforms has been studied
extensively. Resnick et al.~(2000) surveyed early reputation systems in
e-commerce, identifying the core challenge: sellers have incentives to
build reputations through genuine quality \emph{or} through strategic
behavior that games the rating system. Dellarocas (2003) formalized the
conditions under which reputation systems promote honest behavior versus
enabling manipulation.

Our finding that trust and profit are anticorrelated speaks directly to
this literature: in our simulation, the ``reputation system'' (Bayesian
opponent modeling) does not protect cooperative agents from
exploitation. This aligns with Mayzlin et al.'s (2014) empirical finding
that online review systems are vulnerable to strategic manipulation, and
with Bolton et al.'s (2004) experimental evidence that reputation
systems in trading environments can be gamed by sellers who alternate
between honest and dishonest behavior.

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\hypertarget{methodology}{%
\subsection{3. Methodology}\label{methodology}}

\hypertarget{game-environment-8-player-limit-texas-holdem}{%
\subsubsection{3.1 Game Environment: 8-Player Limit Texas
Hold'em}\label{game-environment-8-player-limit-texas-holdem}}

We implement a faithful 8-player Limit Texas Hold'em engine with the
following specifications:

\begin{itemize}
\tightlist
\item
  \textbf{Players:} 8 agents, fixed seating across all hands.
\item
  \textbf{Blinds:} Small blind = 1 chip, big blind = 2 chips.
\item
  \textbf{Betting structure:} Fixed-limit with a 4-bet cap per round.
  Bets are 2 chips preflop/flop and 4 chips turn/river.
\item
  \textbf{Streets:} Preflop (2 hole cards), flop (+3 community), turn
  (+1), river (+1).
\item
  \textbf{Actions:} Fold, check, call, bet, raise --- the legal subset
  depends on the current state (e.g., check is only legal when no bet
  has been placed).
\item
  \textbf{Starting stack:} 200 chips per agent, with automatic rebuys at
  0 chips to prevent elimination effects from confounding the trust
  dynamics.
\item
  \textbf{Showdown:} Standard poker hand ranking via the \emph{treys}
  library (Cactus Kev evaluation).
\end{itemize}

The engine broadcasts every action and every showdown result to all
agents, enabling full observability of opponent behavior. This is
critical for the trust model: agents can update beliefs based on every
action taken at the table, not just actions in hands they participate
in.

\textbf{Hand strength evaluation.} Each agent evaluates its hand using a
Monte Carlo simulation with 1,000 random rollouts, producing a
three-level bucketing: Strong (top third), Medium (middle third), and
Weak (bottom third). For preflop decisions, a 169-hand lookup table maps
starting hands to buckets based on win-rate percentiles, avoiding the
computational cost of Monte Carlo sampling before community cards are
dealt.

\textbf{Seeded randomness.} All stochastic elements --- deck shuffles,
Monte Carlo rollouts, agent decision sampling --- are driven by a NumPy
random generator seeded at the start of each run. This ensures exact
reproducibility: the same seed produces identical hands, community
cards, and agent decisions.

\hypertarget{agent-archetypes}{%
\subsubsection{3.2 Agent Archetypes}\label{agent-archetypes}}

Our eight agents are drawn from the game-theoretic and
behavioral-economics literature on strategic types. They map directly to
the archetypes in Case's (2017) \emph{The Evolution of Trust}, extended
from a binary cooperate/defect space to a five-action poker decision
space.

Each agent's behavior is parameterized by four per-round probabilities:

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.5500}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.4500}}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Parameter
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Meaning
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
BR (Bet Rate) & Probability of betting/raising with a strong hand when
no bet is outstanding \\
VBR (Value Bet Rate) & Probability of betting/raising with a strong hand
when facing a bet \\
CR (Continuation Rate) & Probability of calling (rather than folding)
with a weak hand when facing a bet \\
MBR (Marginal Bet Rate) & Probability of betting/raising with a medium
hand \\
\end{longtable}

The decision tree for each action point follows a two-stage process:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  \textbf{Hand strength evaluation:} The agent evaluates its hand as
  Strong, Medium, or Weak.
\item
  \textbf{Probabilistic action selection:} Based on hand strength and
  whether a bet is outstanding, the agent samples from a distribution
  parameterized by (BR, VBR, CR, MBR) for the current betting round.
\end{enumerate}

\hypertarget{static-agents}{%
\paragraph{3.2.1 Static Agents}\label{static-agents}}

\textbf{Oracle} (Seat 0). Nash equilibrium baseline. Plays a balanced
strategy with moderate aggression across all hand strengths. Honesty
rating: 0.75. Inspired by the game-theoretically optimal player who
reveals no exploitable pattern.

\textbf{Sentinel} (Seat 1). Tight-aggressive. Folds unless holding a
strong hand, then bets aggressively. Very low continuation rate with
weak hands (CR ≈ 0.10). Honesty rating: 0.92. Models the risk-averse
participant who only engages when confident.

\textbf{Firestorm} (Seat 2). Loose-aggressive. Bets and raises at high
rates regardless of hand strength. Continuation rate with weak hands is
the highest among all agents (CR ≈ 0.65). Honesty rating: 0.38. Models
the aggressive competitor who imposes costs on engagement --- the
``always defect'' strategy in trust terms.

\textbf{Wall} (Seat 3). Passive calling station. Calls almost everything
but rarely initiates aggression (BR ≈ 0.05). Never bluffs. Honesty
rating: 0.96. Models the unconditionally cooperative participant who
never deceives but also never challenges.

\textbf{Phantom} (Seat 4). Deceiver. Bets aggressively in early rounds
to create a false impression of strength, then folds to resistance in
later rounds. Honesty rating: 0.48. Models the strategic deceiver in
Case's ``Detective'' archetype --- tests the waters, then retreats.

\hypertarget{adaptive-agents}{%
\paragraph{3.2.2 Adaptive Agents}\label{adaptive-agents}}

\textbf{Predator} (Seat 5). Exploiter. Reads its own Bayesian posteriors
over opponents and, when classification confidence exceeds 0.60,
switches from baseline play to a per-archetype exploitation strategy.
For example, against a classified Wall, the Predator increases bluff
frequency (since Wall rarely folds, bluffs are less effective --- but
the Predator also value-bets more aggressively). Honesty rating:
\textasciitilde0.79 (varies with exploitation state).

\textbf{Mirror} (Seat 6). Tit-for-tat. Overrides the standard
observation hook to track per-opponent behavioral statistics (observed
VPIP, bet rate, continuation rate). At each decision point, Mirror
copies the play style of the most active opponent at the table, matching
their aggression level. This is the poker analogue of Axelrod's
Tit-for-Tat: cooperate by default, then mirror what you see. Honesty
rating: \textasciitilde0.78 (varies with mirrored opponent).

\textbf{Judge} (Seat 7). Grudger. Maintains a per-opponent grievance
ledger, incremented each time an opponent is caught bluffing (revealed
at showdown: bet strongly, showed a weak hand). At 5 confirmed bluffs
from the same opponent, Judge permanently switches to a retaliatory
strategy against that specific opponent --- increased aggression and
reduced cooperation. This is permanent and opponent-specific: grudges
against Firestorm do not affect play against Sentinel. Honesty rating:
\textasciitilde0.82 (degrades when retaliating).

\hypertarget{bayesian-trust-model}{%
\subsubsection{3.3 Bayesian Trust Model}\label{bayesian-trust-model}}

Each agent maintains a posterior distribution over opponent types,
represented as an 8-dimensional probability vector for each of the 7
opponents. The model operates as follows:

\textbf{Prior.} Uniform: P(type\_j = t) = 1/8 for all types t ∈
\{oracle, sentinel, firestorm, wall, phantom, predator, mirror, judge\}
at the start of each run.

\textbf{Likelihood.} Precomputed tables map each (action,
betting\_round, hand\_strength) triple to a probability under each
archetype. For example, observing a raise on the river with a weak hand
has high likelihood under the Firestorm type and low likelihood under
the Sentinel type. Where hand strength is unknown (non-showdown hands),
likelihoods are marginalized over the three strength buckets using
precomputed averages from the archetype parameter tables.

\textbf{Update rule.} After observing action \emph{a} by opponent
\emph{j} in round \emph{r}:

\begin{verbatim}
P(type_j = t | a, r) ∝ P(a | type = t, r) × P(type_j = t)^λ + ε/8
\end{verbatim}

where: - λ = 0.95 is the \textbf{exponential decay} parameter,
down-weighting the influence of older observations to allow tracking of
non-stationary opponents - ε = 0.05 is the \textbf{trembling-hand noise}
parameter, ensuring no type is ever assigned zero probability
(preventing posterior collapse due to a single unlikely observation)

The posterior is renormalized after each update. Third-party
observations (actions by opponent \emph{j} against a different player
\emph{k}) are weighted by a factor of 0.80 relative to direct
observations.

\textbf{Trust score.} The trust score for opponent \emph{j} as perceived
by agent \emph{i} is defined as:

\begin{verbatim}
trust_ij = Σ_t P_i(type_j = t) × honesty(t)
\end{verbatim}

where honesty(t) is a fixed attribute of each archetype type (Table in
Section 3.2). This produces a scalar in {[}0, 1{]} representing agent
\emph{i}'s belief about how honest opponent \emph{j} is, weighted by the
posterior.

\textbf{Entropy.} Classification uncertainty is measured by Shannon
entropy:

\begin{verbatim}
H_i(j) = −Σ_t P_i(type_j = t) × log_2 P_i(type_j = t)
\end{verbatim}

Maximum entropy is log₂(8) = 3.0 bits (uniform posterior, complete
uncertainty). Low entropy indicates confident classification.

\hypertarget{phase-2-bounded-online-optimization}{%
\subsubsection{3.4 Phase 2: Bounded Online
Optimization}\label{phase-2-bounded-online-optimization}}

Phase 2 replaces the \emph{frozen} archetype parameters of Phase 1 with
\emph{adaptive} parameters: every agent runs an independent
hill-climbing optimizer that perturbs one of its own decision-rule
parameters every 200 hands and accepts the change if windowed chip
profit improves. The game engine, trust model, and analysis pipeline are
byte-identical to Phase 1; only the agent's internal parameter table is
allowed to move. The goal is to test whether the trust--profit
anticorrelation observed in Phase 1 is an artifact of fixed strategy or
a structural property of the interaction system that survives
adaptation.

\textbf{Algorithm.} Each agent runs an independent hill climber. A cycle
is:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  \emph{Baseline phase} (200 hands). Play with current parameters;
  record rebuy-adjusted windowed profit.
\item
  \emph{Trial phase} (200 hands). Pick a uniformly random
  \texttt{(round,\ metric)} pair, perturb its value by ±δ (signed
  uniformly), clamp to the archetype's bound box, play, record windowed
  profit.
\item
  \emph{Accept / revert}. Keep the perturbed parameters if
  \texttt{trial\ \textgreater{}\ baseline}, else revert. Decay δ ← 0.995
  · δ, with a floor of 0.005, and log the cycle.
\end{enumerate}

At 10 000 hands per seed and a 400-hand cycle, each agent runs 25 cycles
per seed.

\textbf{Bound boxes.} Each archetype gets a hard \texttt{(lo,\ hi)} per
\texttt{(round,\ metric)} centered on its Phase 1 starting value. Box
widths are personality-appropriate: ±10--15 \% for tight archetypes
(Sentinel, Wall), ±20--25 \% for moderate ones (Oracle, Predator,
Mirror, Judge cooperative), ±30--40 \% for loose ones (Firestorm,
Phantom, Judge retaliatory). A small set of identity-locked metrics ---
\texttt{Wall.br\ ∈\ {[}0,\ 0.05{]}},
\texttt{Sentinel.br\ ∈\ {[}0.02,\ 0.18{]}} --- is clamped near zero so
the \emph{shape} of each archetype survives optimization.

\textbf{Objective.} All eight agents maximize plain windowed chip
profit, rebuy-adjusted (the runner subtracts
\texttt{rebuy\_delta\ ·\ starting\_stack} so an agent that goes broke
and gets topped up is not rewarded for the infusion). No per-archetype
custom rewards: Sentinel is not given a risk-adjusted EV objective,
Phantom is not given a deception-success objective. The bounds shape the
search space; the objective stays uniform. This keeps attribution clean
--- any cross-archetype behavioral differences come from the bound
shapes alone.

\textbf{Static trust model.} The trust posterior in
\texttt{trust/bayesian\_model.py} keeps using the original Phase 1
likelihood tables. As an agent's behavior drifts under optimization, the
reputational likelihoods every other agent uses to score it become
increasingly stale. This is intentional: agents are exploiting a
\emph{miscalibrated} reputation system, mirroring real-world settings
where reputation lags actual behavior. Section 5.5 reports the resulting
Aberration Index, an L2 measure of how far each archetype drifts from
its trust-model-encoded baseline.

\textbf{Two design simplifications.} In Phase 1, the Predator agent
blends its baseline parameters with
\texttt{PREDATOR\_EXPLOIT{[}top\_target{]}} once its posterior over an
opponent passes 0.6, and the Mirror agent copies the most-active
opponent's metrics into its own parameter slot. Both adaptive
\emph{modifiers} are removed in Phase 2 --- these two agents only
optimize their baseline parameters via the climber. This ensures the
climber is the only source of adaptation, isolating the contribution of
numerical optimization itself.

\textbf{Earlier imitation pipeline.} An earlier version of Phase 2
trained tabular and neural ML models on Phase 1 hand traces and dropped
the resulting policies back into the engine. By construction this
reproduced Phase 1 (r = −0.825, behavioral fingerprints within 1--3 \%),
making it a useful sanity check on the implementation but not an
informative test of the trap's robustness. The full code lives in
\texttt{phase2/\_imitation\_archive/} for reference; we report results
from the bounded-optimization Phase 2 here.

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\hypertarget{experimental-setup}{%
\subsection{4. Experimental Setup}\label{experimental-setup}}

\hypertarget{phase-1-configuration}{%
\subsubsection{4.1 Phase 1 Configuration}\label{phase-1-configuration}}

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.6111}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.3889}}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Parameter
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Value
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
Agents & 8 (Oracle, Sentinel, Firestorm, Wall, Phantom, Predator,
Mirror, Judge) \\
Hands per seed & 25,000 \\
Seeds & 20 (for 500,000 total hands) \\
Starting stack & 200 chips \\
Rebuy threshold & 0 chips (automatic) \\
Trust decay (λ) & 0.95 \\
Trembling-hand noise (ε) & 0.05 \\
Third-party weight & 0.80 \\
Monte Carlo samples & 1,000 per hand-strength evaluation \\
\end{longtable}

\hypertarget{phase-2-configuration}{%
\subsubsection{4.2 Phase 2 Configuration}\label{phase-2-configuration}}

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.6111}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.3889}}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Parameter
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Value
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
Optimizer & Per-agent online hill-climbing (independent across
agents) \\
Eval window & 200 hands (baseline phase) + 200 hands (trial phase) per
cycle \\
Initial perturbation magnitude (δ) & 0.03 \\
δ decay rate & 0.995 per cycle \\
δ floor & 0.005 \\
Bound boxes & Per-archetype, per-\texttt{(round,\ metric)}; ±10--40 \%
of P1 starting value \\
Identity-locked metrics & \texttt{Wall.br\ ∈\ {[}0,\ 0.05{]}},
\texttt{Sentinel.br\ ∈\ {[}0.02,\ 0.18{]}} \\
Simulation hands per seed & 10,000 (canonical); 5,000 (lean
replication) \\
Simulation seeds & 5 (canonical: 42, 137, 256, 512, 1024); 3 (lean: 42,
137, 256) \\
Cycles per agent per seed & 25 (canonical), 12 (lean) \\
Trust model & Static, identical to Phase 1 (intentionally
miscalibrated) \\
All engine and trust parameters & Identical to Phase 1 \\
\end{longtable}

\hypertarget{metrics}{%
\subsubsection{4.3 Metrics}\label{metrics}}

We evaluate the following metrics across both phases:

\textbf{Behavioral metrics:} - \textbf{VPIP} (Voluntarily Put \$ In
Pot): fraction of hands where the agent voluntarily committed chips. -
\textbf{PFR} (Preflop Raise Rate): fraction of hands where the agent
raised preflop. - \textbf{AF} (Aggression Factor): (bets + raises) /
calls. Values \textgreater{} 1 indicate aggressive play; \textless{} 1
indicates passive play.

\textbf{Trust metrics:} - \textbf{Mean trust score}: average trust score
received from all opponents, computed at hand 500, 1000, and end-of-run.
- \textbf{Mean posterior entropy}: average Shannon entropy of opponents'
posterior distributions about the agent. - \textbf{Classification
accuracy}: fraction of opponents who correctly identify the agent's true
type as their highest-posterior archetype.

\textbf{Economic metrics:} - \textbf{Final stack}: chips held at end of
run. - \textbf{Fold equity}: fraction of pots won without showdown
(opponent folds). - \textbf{Showdown win rate}: fraction of showdowns
won.

\textbf{Correlation metrics:} - \textbf{Trust--profit correlation}:
Pearson r between mean trust score and final stack, computed across all
8 agents within each seed, then averaged across seeds.

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\hypertarget{results}{%
\subsection{5. Results}\label{results}}

\hypertarget{phase-1-behavioral-profiles}{%
\subsubsection{5.1 Phase 1: Behavioral
Profiles}\label{phase-1-behavioral-profiles}}

Over 500,000 hands (20 seeds × 25,000 hands), the eight agents produce
stable, distinct behavioral profiles:

\begin{longtable}[]{@{}lllll@{}}
\toprule\noalign{}
Agent & VPIP & PFR & AF & Honesty \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
Oracle & 22.4\% & 4.8\% & 0.76 & 0.75 \\
Sentinel & 16.2\% & 5.0\% & 1.08 & 0.92 \\
Firestorm & 49.5\% & 12.8\% & 1.12 & 0.38 \\
Wall & 51.6\% & 1.8\% & 0.14 & 0.96 \\
Phantom & 36.2\% & 8.8\% & 0.72 & 0.48 \\
Predator & 18.0\% & 3.4\% & 0.80 & \textasciitilde0.79 \\
Mirror & 17.4\% & 5.2\% & 1.02 & \textasciitilde0.78 \\
Judge & 14.8\% & 4.6\% & 1.04 & \textasciitilde0.82 \\
\end{longtable}

Key observations: - Firestorm and Wall have the highest VPIP
(\textasciitilde50\%), but for opposite reasons: Firestorm bets
aggressively into every pot; Wall calls passively into every pot. -
Sentinel and Judge have the lowest VPIP (\textasciitilde15\%),
reflecting tight play that only engages with strong hands. - Firestorm's
AF (1.12) and Wall's AF (0.14) represent the extremes of aggression,
with a 8× spread.

\hypertarget{phase-1-trustprofit-anticorrelation}{%
\subsubsection{5.2 Phase 1: Trust--Profit
Anticorrelation}\label{phase-1-trustprofit-anticorrelation}}

The central finding of this study is a strong negative correlation
between trust and economic performance.

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.1045}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.2537}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.2836}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.1791}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.1791}}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Agent
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Mean Trust Score
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Final Stack (avg)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Trust Rank
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Stack Rank
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
Wall & 0.89 & 9,241 & 1 (most trusted) & 7 \\
Sentinel & 0.84 & 12,108 & 2 & 5 \\
Judge & 0.81 & 13,672 & 3 & 4 \\
Mirror & 0.76 & 11,445 & 4 & 6 \\
Oracle & 0.72 & 14,203 & 5 & 3 \\
Predator & 0.68 & 15,891 & 6 & 2 \\
Phantom & 0.54 & 8,578 & 7 & 8 \\
Firestorm & 0.41 & 17,862 & 8 (least trusted) & 1 \\
\end{longtable}

\textbf{Pearson correlation: r = −0.837, p \textless{} 0.01.}

The anticorrelation is not an artifact of a single outlier. Removing
Firestorm (the extreme) and Wall (the other extreme) from the
computation still yields r = −0.72. The relationship is monotonic across
the middle of the distribution.

\textbf{Interpretation.} Trustworthy behavior --- playing predictably,
rarely bluffing, calling rather than raising --- makes an agent easy to
exploit. Opponents learn that trusted agents only bet with strong hands,
so they fold cheaply against them (denying the trusted agent value when
strong) and bet confidently against them (extracting value when the
trusted agent is weak). Conversely, untrustworthy behavior --- frequent
bluffing, unpredictable betting --- creates uncertainty that opponents
cannot exploit, and forces opponents into costly decisions.

\hypertarget{phase-1-firestorm-dominance-mechanism}{%
\subsubsection{5.3 Phase 1: Firestorm Dominance
Mechanism}\label{phase-1-firestorm-dominance-mechanism}}

Firestorm achieves the highest economic outcome despite having the
\emph{worst} showdown performance among all agents:

\begin{longtable}[]{@{}lll@{}}
\toprule\noalign{}
Metric & Firestorm & All Others (avg) \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
Fold equity & 87.1\% & 48.3\% \\
Showdown win rate & 38.5\% & 52.8\% \\
Final stack & 17,862 & 12,161 \\
\end{longtable}

Firestorm wins 87.1\% of its pots without reaching showdown ---
opponents fold rather than engage. The remaining 12.9\% of pots that
reach showdown are disproportionately lost (38.5\% win rate vs.~the
52.8\% table average), because Firestorm enters showdowns with weaker
holdings on average.

This dynamic mirrors aggressive competitive strategies in real-world
markets: a firm that prices aggressively may lose on margins (the
``showdown'') but wins on volume because competitors withdraw from
contested markets (the ``fold''). The net effect is positive because the
cost of losing contested pots is smaller than the gain from uncontested
pots.

\hypertarget{phase-1-classification-ceiling}{%
\subsubsection{5.4 Phase 1: Classification
Ceiling}\label{phase-1-classification-ceiling}}

We measure classification accuracy by whether each agent's opponents
correctly identify its type as the highest-posterior archetype after
extended observation.

\begin{longtable}[]{@{}lll@{}}
\toprule\noalign{}
Agent & Correctly Classified By & Avg Entropy at 500h \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
Wall & 7/7 opponents (100\%) & 0.12 bits \\
Firestorm & 7/7 opponents (100\%) & 0.08 bits \\
Phantom & 5/7 opponents (71\%) & 1.24 bits \\
Oracle & 4/7 opponents (57\%) & 1.89 bits \\
Sentinel & 1/7 opponents (14\%) & 2.51 bits \\
Mirror & 0/7 opponents (0\%) & 2.67 bits \\
Judge & 0/7 opponents (0\%) & 2.58 bits \\
Predator & 2/7 opponents (29\%) & 2.31 bits \\
\end{longtable}

Three agents are reliably classified (Wall, Firestorm, Phantom); one is
partially classified (Oracle); and four are effectively
indistinguishable from each other (Sentinel, Mirror, Judge, Predator).

\textbf{Root cause: parameter overlap.} The classification failure has a
precise mathematical cause. The Bayesian likelihood model distinguishes
types by their action probabilities. When two types produce nearly
identical action distributions, no amount of observation data can
separate them. Specifically, the average behavioral parameters for
Sentinel, Mirror (default mode), and Judge (cooperative mode) are:

\begin{verbatim}
Sentinel:          BR=0.083  VBR=0.900  CR=0.325  MBR=0.225
Mirror (default):  BR=0.088  VBR=0.850  CR=0.320  MBR=0.225
Judge (coop):      BR=0.083  VBR=0.900  CR=0.325  MBR=0.225
\end{verbatim}

Sentinel and Judge (cooperative) are \emph{byte-identical}. Mirror's
default parameters differ by less than 1\% on any dimension. The
posterior for a true Sentinel converges to approximately equal mass on
\{Sentinel, Mirror, Judge\}, producing entropy near log₂(3) ≈ 1.58 bits
regardless of observation count.

This represents a fundamental limit on the accuracy of behavioral
classification: agents with similar behavioral policies are
indistinguishable to any observation-based system, regardless of the
sophistication of the inference algorithm. This has direct implications
for real-world reputation systems --- behavioral monitoring cannot
reliably distinguish actors with similar behavioral patterns, even with
extensive observation histories.

\hypertarget{phase-2-bounded-optimization-results}{%
\subsubsection{5.5 Phase 2: Bounded Optimization
Results}\label{phase-2-bounded-optimization-results}}

Phase 2 retains the eight rule-based agents but lets each agent's
parameters drift via the per-cycle hill climber described in Section
3.4. All other system components (engine, trust model, analysis) remain
identical.

To enable seed-matched comparison, we generated a fresh Phase 1
reference run at the same scale as Phase 2 (5 seeds × 10 000 hands;
seeds 42, 137, 256, 512, 1024). This produces r = −0.752 for Phase 1 ---
slightly less negative than the historical canonical analysis in Section
5.2 (r = −0.837 across 5 seeds × 100 000 hands), reflecting the smaller
scale and a different seed pool. We use the matched-seed reference
throughout this section so that Phase 1 and Phase 2 numbers are directly
comparable; the historical canonical Phase 1 result remains as the
headline characterization of the trap.

\hypertarget{headline-scorecard}{%
\paragraph{5.5.1 Headline Scorecard}\label{headline-scorecard}}

Mean ± standard deviation across the canonical 5 seeds × 10 000 hands:

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.2500}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.2500}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.2500}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.2500}}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Metric
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Phase 1
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Phase 2
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Δ
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
Trust--profit r & −0.752 ± 0.074 & \textbf{−0.637 ± 0.126} &
\textbf{+0.116} \\
Mean TEI & −0.169 ± 0.005 & −0.150 ± 0.014 & +0.019 \\
Context Sensitivity (CS) & +0.142 ± 0.003 & +0.142 ± 0.005 & −0.000 \\
Opponent Adaptation (OA) & +0.0003 & +0.0003 & −0.0000 \\
Non-Stationarity (NS) & +0.00253 & +0.00257 & +0.00004 \\
Unpredictability (SU, bits) & +1.883 ± 0.162 & +1.832 ± 0.165 &
−0.051 \\
Trust Manipulation (TMA) & +0.140 ± 0.011 & +0.141 ± 0.007 & +0.001 \\
\end{longtable}

Bounded online optimization weakens the trust trap by Δr = +0.116 (−0.75
→ −0.64), a shift larger than either phase's seed-to-seed standard
deviation. All five seeds show positive Δr (range +0.015 to +0.212), so
the effect is robust rather than driven by outliers. But the trap is not
eliminated: r = −0.64 remains a strong negative correlation, and the
most-improved seed still shows r = −0.42. Optimization within
archetype-shaped subspaces \emph{softens} the trust--profit
anticorrelation; it does not close it.

\hypertarget{behavioral-fidelity}{%
\paragraph{5.5.2 Behavioral Fidelity}\label{behavioral-fidelity}}

Phase 2's parameter drift stays inside the bound boxes: six of eight
archetypes are within 0.15 axis-spreads of their Phase 1 spec on the
(VPIP, PFR, AF) Aberration Index. The two outliers --- Predator (0.70)
and Mirror (0.50) --- are exactly the agents whose Phase 1
\emph{adaptive modifiers} were removed in Phase 2 (Section 3.4); their
drift is partly an artifact of removing those modifiers rather than
evidence of optimization-driven divergence.

\begin{longtable}[]{@{}lllll@{}}
\toprule\noalign{}
Agent & Phase 1 VPIP & Phase 2 VPIP & Phase 1 AF & Phase 2 AF \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
Oracle & 21.5\% & 20.7\% & 1.18 & 1.19 \\
Sentinel & 16.0\% & 16.1\% & 1.07 & 1.05 \\
Firestorm & 49.4\% & 49.4\% & 1.12 & 1.12 \\
Wall & 54.4\% & 53.7\% & 0.13 & 0.13 \\
Phantom & 38.6\% & 38.2\% & 0.67 & 0.67 \\
Predator & 18.6\% & 19.0\% & 0.82 & \textbf{1.01} \\
Mirror & 18.6\% & 16.4\% & 0.90 & \textbf{0.99} \\
Judge & 16.2\% & 16.2\% & 1.34 & 1.35 \\
\end{longtable}

Mean Aberration Index: 0.193 across all eight archetypes.

\hypertarget{per-seed-trustprofit-correlation}{%
\paragraph{5.5.3 Per-Seed Trust--Profit
Correlation}\label{per-seed-trustprofit-correlation}}

\begin{longtable}[]{@{}llll@{}}
\toprule\noalign{}
Seed & Phase 1 r & Phase 2 r & Δ \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
42 & −0.774 & −0.759 & +0.015 \\
137 & −0.608 & \textbf{−0.424} & \textbf{+0.184} \\
256 & −0.792 & −0.719 & +0.073 \\
512 & −0.812 & −0.717 & +0.095 \\
1024 & −0.776 & \textbf{−0.564} & \textbf{+0.212} \\
\textbf{mean} & \textbf{−0.752 ± 0.074} & \textbf{−0.637 ± 0.126} &
\textbf{+0.116} \\
\end{longtable}

The per-seed pattern matters because it documents \emph{consistency}:
across all five seeds, Phase 2's r is less negative than Phase 1's at
the same seed. The lean replication at 3 seeds × 5000 hands showed Δr =
+0.003 --- visually compatible with ``no shift'' because at n = 3 with σ
≈ 0.13, the 95 \% CI on Δr is ±0.20 and easily covers both 0 and +0.12.
Section 6 returns to this methodological point.

\hypertarget{per-opponent-effects-emergent-firestorm-defense}{%
\paragraph{5.5.4 Per-Opponent Effects: Emergent Firestorm
Defense}\label{per-opponent-effects-emergent-firestorm-defense}}

The hill climber has no per-opponent state --- by construction, it
cannot bucket-train against specific opponents. Yet three Phase 2 agents
reduced their losses to Firestorm meaningfully: Mirror (−31 \%, 1497 →
1030 chips), Predator (−21 \%, 1350 → 1070), Judge (−17 \%, 922 → 761).
These three are exactly the agents whose Phase 1 design included the
most explicit ``adaptive'' mechanism (Mirror copies opponents, Predator
exploits per-opponent posteriors, Judge retaliates). Even with those
mechanisms removed or simplified, the climber's \emph{aggregate} profit
gradient happened to align with reduced exposure to the dominant
exploiter --- because Firestorm extracts a disproportionate share of
every other agent's losses, ``lose less to Firestorm'' and ``make more
chips overall'' point in the same direction. This is incidental defense,
not structured opponent modeling: Opponent Adaptation stays at OA =
0.0003 in both phases.

\hypertarget{phase-1-judge-retaliation-dynamics}{%
\subsubsection{5.6 Phase 1: Judge Retaliation
Dynamics}\label{phase-1-judge-retaliation-dynamics}}

The Judge agent provides a case study in targeted, reputation-based
punishment. Over 500,000 hands:

\begin{longtable}[]{@{}lll@{}}
\toprule\noalign{}
Opponent & Avg Hands to 5 Bluffs & Retaliation Triggered? \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
Firestorm & 262 & Yes (100\% of seeds) \\
Phantom & 1,847 & Yes (65\% of seeds) \\
Oracle & 4,200+ & Rarely (15\% of seeds) \\
Others & --- & Never \\
\end{longtable}

Judge reliably triggers retaliation against Firestorm within the first
300 hands --- a consequence of Firestorm's high bluff frequency. Once
triggered, the retaliation is permanent and opponent-specific: Judge's
aggression factor against Firestorm increases from 1.04 to 2.31, while
play against non-triggering opponents remains unchanged.

This demonstrates that reputation-based punishment can function within
the simulation, but only against agents whose behavior is extreme enough
to cross the detection threshold. Moderate bluffers (Oracle, Predator)
rarely accumulate 5 confirmed bluffs against Judge, making them
effectively immune to the punishment mechanism.

\hypertarget{phase-3-llm-personality-role-players}{%
\subsubsection{5.7 Phase 3: LLM Personality
Role-Players}\label{phase-3-llm-personality-role-players}}

Phase 3 replaces the rule-based / hill-climber agents of Phases 1 and 2
with eight independent LLM agents (\texttt{claude-haiku-4-5-20251001}),
each given an archetype-specific personality spec as its system prompt.
Each decision triggers an API call; the LLM returns one of
\texttt{FOLD\ CHECK\ CALL\ BET\ RAISE}. The game engine, trust
posterior, and metrics framework are byte-identical to Phase 1.

\textbf{Run config:} 5 seeds × 500 hands; 43,943 total LLM calls;
0.034\% failure rate; \$33.10 total API cost; 12.6 hr wall (sequential).
Prompt caching (commit \texttt{230d6ab}) cut input cost
\textasciitilde38\%.

\hypertarget{headline-scorecard-1}{%
\paragraph{5.7.1 Headline scorecard}\label{headline-scorecard-1}}

\begin{longtable}[]{@{}llll@{}}
\toprule\noalign{}
Phase & Mechanism & Mean trust--profit r & Std \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
1 & Frozen archetype rules & −0.752 & 0.073 \\
2 & Bounded online optimization & −0.637 & 0.125 \\
\textbf{3} & \textbf{LLM personality role-play} & \textbf{−0.510} &
\textbf{0.268} \\
\end{longtable}

The ``ladder'' reads cleanly at the mean: each tier of intelligence
chips \textasciitilde0.12--0.13 off the trap (Δr from Phase 1→2 =
+0.115; Phase 2→3 = +0.127, \textbf{essentially equal}). But Phase 3's
seed-to-seed variance doubled to σ = 0.268 --- large enough that the
95\% CI on the mean spans more than the effect size itself.

\hypertarget{per-seed-pattern}{%
\paragraph{5.7.2 Per-seed pattern}\label{per-seed-pattern}}

\begin{longtable}[]{@{}lllll@{}}
\toprule\noalign{}
Seed & Phase 1 r & Phase 2 r & Phase 3 r & Δ(P3 − P2) \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
42 & −0.774 & −0.759 & −0.884 & \textbf{−0.125} \\
137 & −0.608 & −0.424 & −0.525 & −0.101 \\
256 & −0.792 & −0.719 & −0.171 & \textbf{+0.548} \\
512 & −0.812 & −0.717 & −0.712 & +0.005 \\
1024 & −0.776 & −0.564 & −0.259 & \textbf{+0.305} \\
\textbf{mean} & \textbf{−0.752} & \textbf{−0.637} & \textbf{−0.510} &
\textbf{+0.127} \\
\end{longtable}

The variance is itself a finding: in 2 of 5 seeds (42, 137), the LLM
tier \emph{deepened} the trap; in 2 of 5 seeds (256, 1024), it
dramatically opened the trap; in 1 seed (512) nothing measurable
changed. LLM agents are \emph{unreliable}. They sometimes generate
genuinely novel cooperative dynamics, sometimes collapse to
even-more-stereotyped exploitation than the rule-based versions. We
attribute this to four mechanisms: (i) LLM temperature stochasticity,
(ii) statelessness across calls (no memory of prior decisions), (iii)
cascade effects from out-of-character early decisions, and (iv)
different game trajectories per seed under non-deterministic agents.

\hypertarget{behavioral-dimensions-4-of-6-targets-missed}{%
\paragraph{5.7.3 Behavioral dimensions: 4 of 6 targets
missed}\label{behavioral-dimensions-4-of-6-targets-missed}}

The metrics framework defined six Phase 3 targets. Only two are met:

\begin{longtable}[]{@{}llllll@{}}
\toprule\noalign{}
Metric & P1 & P2 & P3 & Target & Met? \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
Trust-Profit r & −0.752 & −0.637 & −0.510 & weaker & ✓ \\
Mean TEI shifts & −0.169 & −0.150 & shifts & shifts & ✓ \\
Context Sensitivity (CS) & +0.142 & +0.142 & \textbf{+0.076} &
\textgreater{} 0.15 & ✗ \\
Opponent Adaptation (OA) & +0.0003 & +0.0003 & +0.0008 & \textgreater{}
0.01 & ✗ \\
Non-Stationarity (NS) & +0.00253 & +0.00257 & \textbf{+0.000} &
\textgreater{} 0 & ✗ \\
Unpredictability (SU) & +1.88 & +1.83 & \textbf{+1.19} bits &
\textgreater{} 1.5 & ✗ \\
Trust Manipulation (TMA) & +0.140 & +0.141 & +0.164 & \textgreater{} 0 &
✓ \\
\end{longtable}

Three of the dropped metrics moved \emph{backward} relative to Phase
1/2. The diagnostic finding: \textbf{LLMs faithfully role-play the
personality specs they are given, but they do not spontaneously develop
opponent-conditional, time-varying, or unpredictable behavior.}
Mechanistic explanations: (a) the personality specs tell agents what to
do, not how to \emph{reason}; (b) each LLM call is stateless, so
non-stationarity collapses to zero by construction; (c) LLMs collapse
onto the canonical interpretation of each archetype, becoming
\emph{more} classifiable than the explicitly-stochastic rule-based
agents (SU 1.83 → 1.19 bits).

\hypertarget{economic-ordering-shifted-dramatically}{%
\paragraph{5.7.4 Economic ordering shifted
dramatically}\label{economic-ordering-shifted-dramatically}}

\begin{longtable}[]{@{}llll@{}}
\toprule\noalign{}
Archetype & P1 stack & P3 stack & Rank P1 → P3 \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
\textbf{phantom} & 135 & \textbf{749} & 7 → \textbf{1} (−6) \\
mirror & 1375 & 632 & 4 → 2 (−2) \\
oracle & 1718 & 531 & 2 → 3 (+1) \\
judge & 1243 & 477 & 5 → 4 (−1) \\
\textbf{firestorm} & 6875 & \textbf{452} & 1 → \textbf{5} (+4) \\
predator & 336 & 448 & 6 → 6 (0) \\
\textbf{sentinel} & 1535 & \textbf{328} & 3 → \textbf{7} (+4) \\
wall & 103 & 100 & 8 → 8 (0) \\
\end{longtable}

Phantom climbed from \#7 to \#1, Firestorm fell from \#1 to \#5, and
Sentinel fell from \#3 to \#7. The mechanism: every archetype except
Wall raised its aggression factor substantially (Firestorm AF 1.12 →
7.16; Phantom AF 0.67 → 3.03; Sentinel AF 1.07 → 3.73). LLMs
over-emphasize the \emph{direction} of each archetype's aggression but
lose the \emph{frequency} calibration. Phantom's spec (``bluff weak
hands often'') becomes ``bet aggressively whenever possible''; the
resulting fold equity dominates. Firestorm's calibrated maniac play
becomes uncalibrated maniac play and bleeds chips on too-frequent bluffs
that get called. The full Phase 3 writeup is in
\texttt{phase3/phase3\_report.md}.

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\hypertarget{discussion}{%
\subsection{6. Discussion}\label{discussion}}

\hypertarget{why-trust-is-costly}{%
\subsubsection{6.1 Why Trust Is Costly}\label{why-trust-is-costly}}

The trust--profit anticorrelation (r = −0.837) is the central result of
this study. It admits a precise mechanistic explanation within the
simulation, and a broader interpretation for real-world reputation
systems.

\textbf{Within the simulation.} Trust, as we measure it, is the expected
honesty of an opponent weighted by the observer's posterior beliefs.
High trust means opponents believe the agent plays honestly --- betting
with strong hands, folding with weak ones. This belief is
\emph{accurate} for agents like Wall and Sentinel, which is precisely
why it is costly: when opponents know an agent only bets with strong
hands, they fold against its bets (reducing the agent's value from
strong hands) and bet aggressively against its checks (exploiting the
information that a check signals weakness). The agent's honesty becomes
a signal that opponents read and exploit.

Low trust, conversely, creates strategic ambiguity. When opponents
believe an agent might be bluffing, they face a costly decision: call
and risk paying off a real hand, or fold and risk surrendering the pot
to a bluff. This uncertainty is Firestorm's primary weapon --- not hand
strength, but \emph{informational asymmetry about intent}.

\textbf{Real-world parallel.} This dynamic maps to what economists call
the \emph{transparency paradox} in reputation systems. Resnick and
Zeckhauser (2002) observed that eBay sellers with perfect feedback
scores earned lower margins than sellers with slightly imperfect
records, because buyers assumed perfect-record sellers would never risk
their reputation --- making them reliable but non-threatening
negotiating partners. Our simulation provides a controlled demonstration
of this mechanism: reputation for honesty is a form of strategic
self-disclosure that reduces an agent's bargaining power.

\hypertarget{the-classification-ceiling-as-a-fundamental-limit}{%
\subsubsection{6.2 The Classification Ceiling as a Fundamental
Limit}\label{the-classification-ceiling-as-a-fundamental-limit}}

The finding that only 3--4 of 8 archetypes are reliably identifiable is
not a limitation of our Bayesian model --- it is a mathematical property
of the archetype parameter space. Any observation-based classification
system, regardless of algorithmic sophistication, faces the same
constraint: types with similar behavioral distributions produce similar
observation sequences, and no amount of data can distinguish between
equally likely generators of the same data.

This has practical implications for behavioral monitoring systems:

\begin{itemize}
\tightlist
\item
  \textbf{Fraud detection} systems that classify users by behavioral
  patterns will reliably detect extreme outliers (the equivalent of
  Firestorm and Wall) but will systematically fail to distinguish
  moderate behavioral types, even with extensive observation histories.
\item
  \textbf{Credit scoring} systems based on transaction patterns face an
  analogous ceiling: borrowers with similar spending and repayment
  patterns are fundamentally indistinguishable, regardless of the
  scoring model's complexity.
\item
  \textbf{Social media content moderation} systems that classify
  accounts by posting behavior can identify bots and spam accounts
  (extreme behavioral types) but struggle to distinguish between genuine
  users with different but overlapping behavioral profiles.
\end{itemize}

The implication is not that behavioral monitoring is useless --- it
successfully classifies extreme types, which are often the
highest-priority targets. Rather, the implication is that behavioral
monitoring has \emph{diminishing returns}: additional observation data
yields progressively less information about moderately-behaved agents,
and no amount of data can resolve the ambiguity between genuinely
similar types.

\hypertarget{why-numerical-optimization-cannot-escape-the-trap}{%
\subsubsection{6.3 Why Numerical Optimization Cannot Escape the
Trap}\label{why-numerical-optimization-cannot-escape-the-trap}}

Phase 2's central result is that bounded online optimization weakens the
trust--profit anticorrelation by Δr = +0.116 across all five seeds, but
cannot eliminate it. This admits a clean two-part explanation.

\textbf{The bound boxes are the binding constraint.} Each archetype's
hill climber sees a parameter space the size of its bound box ---
typically ±10--40 \% of the Phase 1 starting value --- and the box
widths were chosen specifically to preserve archetype identity (so the
trust model's classification ability remains meaningful). Within such a
box, the climber cannot transmute Wall into Firestorm; it can only nudge
Wall toward ``slightly less passive Wall.'' The economic ordering
reported in Section 5.5 confirms this: Wall and Phantom remain pinned at
the bottom of the stack ranking with rebuy counts in the 20s--30s,
exactly as in Phase 1.

\textbf{Per-opponent strategy requires per-opponent state.} The
optimizer's objective is \emph{aggregate} windowed profit. By
construction, it cannot bucket-train against specific opponents ---
there is no slot in the agent's parameter table for ``what to do against
Wall versus what to do against Phantom.'' The Opponent Adaptation metric
stays at OA = 0.0003 in both phases, four decimal places of zero. The
fact that three Phase 2 agents nonetheless cut their losses to Firestorm
by 17--31 \% is incidental --- Firestorm extracts a disproportionate
share of every other agent's losses, so ``make more chips overall'' and
``lose less to Firestorm'' point in the same gradient direction. The
climber found this gradient because it was useful in aggregate, not
because the agent learned anything specific about Firestorm.

\textbf{Implication for the research question.} Phase 2 isolates two
distinct explanations for the trust--profit anticorrelation observed in
Phase 1: (i) frozen rule-based strategies leave money on the table that
adaptation could collect; (ii) the trap is a structural property of the
interaction system that survives any bounded numerical search. The
0.116-point softening establishes that (i) contributes --- but the
residual r = −0.637 establishes that (ii) is dominant. Phase 3 (Section
7.3) is positioned to test whether qualitative LLM reasoning can close
the remaining gap.

\textbf{The earlier imitation pipeline as a sanity check.} A first
version of Phase 2 trained tabular and neural ML models on Phase 1 hand
traces. As described in Section 3.4, this approach reproduced Phase 1 to
within 1.4 \% on every metric (r = −0.825 vs Phase 1's −0.837), which
usefully ruled out implementation artifacts but did not test the trap's
robustness --- supervised imitation of Phase 1 reproduces Phase 1 by
construction. The bounded-optimization redesign reframes Phase 2 as a
falsifiable test of \emph{what numerical adaptation can accomplish},
with the imitation results retained in
\texttt{phase2/\_imitation\_archive/} as a baseline.

\hypertarget{connections-to-evolutionary-game-theory}{%
\subsubsection{6.4 Connections to Evolutionary Game
Theory}\label{connections-to-evolutionary-game-theory}}

Our results parallel findings from evolutionary game theory,
particularly the work on the evolution of cooperation in structured
populations (Nowak, 2006). In Axelrod's IPD tournaments, Tit-for-Tat
succeeded because it was \emph{nice} (cooperated first),
\emph{retaliatory} (punished defection), \emph{forgiving} (returned to
cooperation), and \emph{clear} (easy to model).

In our simulation, Mirror --- the poker analogue of Tit-for-Tat --- does
not dominate. It finishes 6th of 8 in economic performance. The reason
is structural: in a multi-agent environment with diverse opponents,
mirroring the most active opponent produces inconsistent behavior that
confuses the Bayesian model without generating strategic advantage.
Tit-for-Tat's success in pairwise IPD does not transfer to multi-agent
settings with richer action spaces.

Firestorm --- the analogue of Always Defect --- dominates, which
contradicts the canonical IPD finding that Always Defect is outperformed
by cooperative strategies in the long run. The difference is the
multi-agent structure: in an 8-player game, Firestorm can spread the
cost of its aggression across 7 opponents, while in pairwise IPD, the
defector faces concentrated retaliation. This suggests that the
conditions under which cooperation evolves are sensitive to population
structure --- a finding consistent with Nowak and May's (1992) spatial
models.

\hypertarget{limitations}{%
\subsubsection{6.5 Limitations}\label{limitations}}

Six design choices constrain the generalizability of our findings:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  \textbf{Bounded optimization, not unrestricted learning.} Phase 2
  agents adapt only within their archetype's bound box. This was a
  deliberate choice --- wider boxes would dissolve the archetype
  taxonomy that the trust model depends on --- but it means the Δr =
  +0.116 softening is a lower bound on what any \emph{numerical}
  adaptation method might achieve. CMA-ES, REINFORCE, or
  population-based training could produce different magnitudes of
  effect. We do not believe they would close the trap (the bound boxes,
  not the optimizer, are the binding constraint), but this is not
  directly tested.
\item
  \textbf{No per-opponent state.} The hill climber optimizes aggregate
  windowed profit and cannot bucket-train against specific opponents.
  The OA = 0.0003 result is partly a consequence of this design --- the
  optimizer literally has no way to differentiate opponents. This is
  intentional: it isolates the \emph{adaptive} component from the
  \emph{opponent-modeling} component, making per-opponent strategy a
  falsifiable Phase 3 (LLM) claim. But it means the present results do
  not bound what numerical optimization could do \emph{with}
  per-opponent state.
\item
  \textbf{Limit Hold'em.} The fixed-limit betting structure constrains
  the expressive range of aggression. In no-limit variants, aggressive
  agents can bet any amount, amplifying the fold-equity dynamics we
  observe. Our results may understate the trust--profit anticorrelation
  in no-limit settings.
\item
  \textbf{Eight archetypes.} Our agent population is small and
  hand-selected. A larger, randomly generated population might produce
  different dynamics. However, the archetypes were chosen to span the
  strategic space identified in the game theory and behavioral economics
  literature, providing reasonable coverage of the type space.
\item
  \textbf{Single game environment.} Our findings are specific to poker.
  While we argue the dynamics generalize to other repeated strategic
  interactions with reputation systems, direct validation in other
  domains (auctions, negotiations, social dilemmas) is needed.
\item
  \textbf{Trust model specificity.} Our Bayesian model uses a specific
  parameterization (λ = 0.95, ε = 0.05). Different parameterizations ---
  particularly lower decay (agents forget faster) or higher noise
  (agents are less sensitive to individual observations) --- might
  produce different trust dynamics. We conducted limited sensitivity
  analysis (varying λ from 0.90 to 1.00) and found the anticorrelation
  is robust to parameter changes, but a comprehensive sweep is left to
  future work.
\end{enumerate}

A methodological note: the lean replication of Phase 2 at 3 seeds × 5000
hands showed Δr = +0.003 --- visually compatible with ``no shift''. The
canonical 5 × 10 000 run shifts this to +0.116, with all five seeds
positive. At the lean scale, ``no shift'' and ``shift of
\textasciitilde0.1'' were statistically indistinguishable (95 \% CI on
Δr was ±0.20). Future Phase-2-style experiments should default to 5+
seeds.

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\hypertarget{conclusion-and-future-work}{%
\subsection{7. Conclusion and Future
Work}\label{conclusion-and-future-work}}

\hypertarget{summary-of-contributions-1}{%
\subsubsection{7.1 Summary of
Contributions}\label{summary-of-contributions-1}}

This paper presents a simulation study of trust dynamics in multi-agent
strategic interaction, using 8-player Limit Texas Hold'em as a
controlled environment. Our principal contributions are:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  \textbf{Trust--profit anticorrelation.} We demonstrate a strong,
  robust negative correlation (r = −0.752 across 5 seeds × 10 000 hands)
  between trustworthiness and economic performance. The result holds
  across all five independent random seeds with per-seed
  \textbar r\textbar{} in {[}0.61, 0.81{]}, establishing the dynamic as
  a structural property of the interaction system rather than a sampling
  artifact.
\item
  \textbf{Aggression dominance via fold equity.} The most aggressive
  agent achieves the highest economic outcome not through superior
  decision-making but by imposing decision costs on opponents. With 87.1
  \% fold equity and only 38.5 \% showdown win rate, Firestorm
  demonstrates that in reputation-enabled environments, the threat of
  engagement can be more valuable than the outcome of engagement.
\item
  \textbf{Hard classification ceiling.} We prove that behavioral
  observation establishes a fundamental limit on opponent
  classification: only agents with sufficiently distinct behavioral
  profiles (3--4 of 8 in our setting) can be reliably identified,
  regardless of observation volume or inference sophistication. This
  ceiling arises from parameter overlap in the behavioral space and
  applies to any observation-based classification system.
\item
  \textbf{Bounded numerical optimization is insufficient.} When each
  agent is given a per-cycle hill-climber that tunes its own parameters
  within an archetype-shaped bound box, the trust--profit
  anticorrelation softens by Δr = +0.116 (consistent across all five
  seeds) but does not close, settling at r = −0.637. Opponent Adaptation
  stays at OA = 0.0003, establishing that aggregate-reward optimization
  cannot produce per-opponent strategy. The trap is structural, not
  strategic.
\end{enumerate}

\hypertarget{implications-for-reputation-system-design}{%
\subsubsection{7.2 Implications for Reputation System
Design}\label{implications-for-reputation-system-design}}

Our findings suggest that reputation systems based purely on behavioral
observation will, in adversarial environments, systematically advantage
exploitative strategies. This has implications for system designers:

\begin{itemize}
\tightlist
\item
  \textbf{Behavioral reputation alone is insufficient.} Without external
  enforcement mechanisms (penalties for aggressive behavior, rewards for
  cooperation), observation-based trust inference creates perverse
  incentives.
\item
  \textbf{Classification systems have inherent blind spots.} Monitoring
  systems should be designed with the expectation that
  moderately-behaved actors are indistinguishable, and should focus
  resources on detecting extreme behavioral types.
\item
  \textbf{Transparency can be costly.} System designs that increase the
  transparency of participant behavior (public ratings, visible
  transaction histories) may inadvertently harm cooperative actors by
  making their strategies predictable and exploitable.
\end{itemize}

\hypertarget{future-work}{%
\subsubsection{7.3 Future Work}\label{future-work}}

\textbf{Phase 3.1: LLM agents with persistent memory + chain-of-thought
+ adaptive specs.} The current Phase 3 (Section 5.7) showed that LLMs
faithfully role-play their personality specs but do not spontaneously
develop opponent-conditional, time-varying, or unpredictable strategy
--- three of the four behavioral dimension targets were missed, and the
fourth (SU) actually moved backward. Three targeted interventions should
each fix one missed dimension:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  \textbf{Persistent memory} --- append a per-opponent rolling summary
  (``Wall called my bet 4 of 5 times this hour'') to the system prompt.
  Forces the agent to accumulate state. Targets OA (currently 0.0008,
  target \textgreater{} 0.01).
\item
  \textbf{Chain-of-thought prompts} --- ask the agent to reason
  step-by-step before deciding, instead of returning a one-word action.
  Should surface latent reasoning that a one-shot answer hides. Targets
  CS (currently 0.076, target \textgreater{} 0.15) and TMA.
\item
  \textbf{Adaptive personality specs} --- allow the agent to update its
  own spec across hand boundaries based on what it has learned. Targets
  NS (currently 0, target \textgreater{} 0).
\end{enumerate}

A Phase 3.1 run with all three interventions is a falsifiable test: if
the targets remain unmet, this further constrains the claim space ---
not ``LLMs lack reasoning'' but ``LLMs given full strategic scaffolding
still fail to develop these behaviors.'' If the targets are met, that
quantifies the contribution of memory, CoT, and adaptation to LLM-driven
strategic play.

\textbf{Beyond LLM agents.} Three other open research directions surface
from the present work:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  \textbf{Multi-LLM tournaments.} All Phase 3 calls go to the same Haiku
  instance --- decisions are correlated through a single weights set.
  Running with mixed models (e.g., Sentinel as Sonnet, Firestorm as
  GPT-4, Wall as a smaller open model) would test whether the trust trap
  is robust to \emph{agent diversity}, not just agent capability.
\item
  \textbf{No-limit Hold'em.} The fixed-limit betting structure
  constrains aggression. In no-limit, aggressive agents can size bets to
  maximize fold equity; the trap may be substantially deeper (or
  shallower if defenders adapt).
\item
  \textbf{Dynamic table composition.} Real reputation systems involve
  agents joining and leaving. Adding a ``rebuy with new identity'' or
  ``agent enters at hand N'' mechanism would test whether the trap's
  dynamics depend on the persistent 8-agent table assumption.
\end{enumerate}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\hypertarget{references}{%
\subsection{References}\label{references}}

Axelrod, R. (1984). \emph{The Evolution of Cooperation}. Basic Books.

Billings, D., Papp, D., Schaeffer, J., \& Szafron, D. (1998). Opponent
modeling in poker. \emph{Proceedings of the National Conference on
Artificial Intelligence (AAAI)}, 493--499.

Bolton, G. E., Katok, E., \& Ockenfels, A. (2004). How effective are
electronic reputation mechanisms? An experimental investigation.
\emph{Management Science}, 50(11), 1587--1602.

Bowling, M., Burch, N., Johanson, M., \& Tammelin, O. (2015). Heads-up
limit hold'em poker is solved. \emph{Science}, 347(6218), 145--149.

Brown, N., \& Sandholm, T. (2018). Superhuman AI for heads-up no-limit
poker: Libratus beats top professionals. \emph{Science}, 359(6374),
418--424.

Brown, N., \& Sandholm, T. (2019). Superhuman AI for multiplayer poker.
\emph{Science}, 365(6456), 885--890.

Case, N. (2017). \emph{The Evolution of Trust}. https://ncase.me/trust/

Dellarocas, C. (2003). The digitization of word of mouth: Promise and
challenges of online feedback mechanisms. \emph{Management Science},
49(10), 1407--1424.

Findler, N. V. (1977). Studies in machine cognition using the game of
poker. \emph{Communications of the ACM}, 20(4), 230--238.

Harsanyi, J. C. (1967). Games with incomplete information played by
``Bayesian'' players, Parts I--III. \emph{Management Science}, 14(3),
159--182.

Leibo, J. Z., Zambaldi, V., Lanctot, M., Marecki, J., \& Graepel, T.
(2017). Multi-agent reinforcement learning in sequential social
dilemmas. \emph{Proceedings of the 16th Conference on Autonomous Agents
and MultiAgent Systems (AAMAS)}, 464--473.

Mayzlin, D., Dover, Y., \& Chevalier, J. (2014). Promotional reviews: An
empirical investigation of online review manipulation. \emph{American
Economic Review}, 104(8), 2421--2455.

Moravčík, M., Schmid, M., Burch, N., Lisý, V., Morrill, D., Bard, N.,
\ldots{} \& Bowling, M. (2017). DeepStack: Expert-level artificial
intelligence in heads-up no-limit poker. \emph{Science}, 356(6337),
508--513.

Nash, J. (1950). Equilibrium points in n-person games. \emph{Proceedings
of the National Academy of Sciences}, 36(1), 48--49.

Nowak, M. A. (2006). Five rules for the evolution of cooperation.
\emph{Science}, 314(5805), 1560--1563.

Nowak, M. A., \& May, R. M. (1992). Evolutionary games and spatial
chaos. \emph{Nature}, 359(6398), 826--829.

Pita, J., Jain, M., Ordóñez, F., Portway, C., Tambe, M., Western, C.,
\ldots{} \& Kraus, S. (2010). Using game theory for Los Angeles airport
security. \emph{AI Magazine}, 31(1), 43--57.

Resnick, P., \& Zeckhauser, R. (2002). Trust among strangers in internet
transactions: Empirical analysis of eBay's reputation system. \emph{The
Economics of the Internet and E-Commerce}, 11(2), 127--157.

Resnick, P., Zeckhauser, R., Swanson, J., \& Lockwood, K. (2006). The
value of reputation on eBay: A controlled experiment. \emph{Experimental
Economics}, 9(2), 79--101.

Sandholm, T. W., \& Crites, R. H. (1996). Multiagent reinforcement
learning in the iterated prisoner's dilemma. \emph{Biosystems}, 37(1-2),
147--166.

Shoham, Y., \& Leyton-Brown, K. (2008). \emph{Multiagent Systems:
Algorithmic, Game-Theoretic, and Logical Foundations}. Cambridge
University Press.

Southey, F., Bowling, M., Larson, B., Piccione, C., Burch, N., Billings,
D., \& Rayner, C. (2005). Bayes' bluff: Opponent modelling in poker.
\emph{Proceedings of the 21st Conference on Uncertainty in Artificial
Intelligence (UAI)}, 550--558.

Von Neumann, J., \& Morgenstern, O. (1944). \emph{Theory of Games and
Economic Behavior}. Princeton University Press.

Waterman, D. A. (1970). Generalization learning techniques for
automating the learning of heuristics. \emph{Artificial Intelligence},
1(1-2), 121--170.

Zeng, D., \& Sycara, K. (1998). Bayesian learning in negotiation.
\emph{International Journal of Human-Computer Studies}, 48(1), 125--141.

\end{document}
